### 2022-02-07 outputs/paper/paper.Rmd
The topic of paper chosen, cleaned dataset and completed data section of the paper.

---
title: "Accessment of Homelessness in Toronto in regards to COVID-19 and other factors"
subtitle: "Data from January 2020 to December 2021"
author: "Ziyao Han"
thanks: "Code and data are available at: https://github.com/HanFrank/STA304-Paper-1"
date: "February 6, 2022"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output: 
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(knitr)
library(bookdown)
library(kableExtra)
library(ggplot2)

# Grabbing dataset
monthly_shelter_usage <- read_csv(here::here("inputs/data/shelter_flow_data.csv"))

# removing id generated by read_csv
monthly_shelter_usage <- monthly_shelter_usage[-c(1)]
```

# Introduction

First Paragraph is going to be motivational and broad.

Second Paragraph is about what was done and what was found.

Third paragraph is about implications.

Final Paragraph about the remainder of the paper.

The statistical analysis in this report will be done using `R` [@citeR]. The R Packages, `tidyverse` [@citetidy] and `dplyr` [@citedplyr] will be used for data manipulation and cleaning. The graphs and tables for this report will created and formatted with `ggplot2` [@citeggplot] and `kableExtra` [@citekable]. The packages `bookdown` [@citebookdown] and `knitr` [@citeknitr] will be used to format this report.

# Data

## Data Source

The data that was used for this report was obtained and gathered by the Shelter, Support & Housing Administration (SSHA) of the City of Toronto [@citedata]. This data is collected through the City of Toronto funded homelessness shelters and services. This data shares information about the people accessing the services and is meant to provide insight regarding homelessness in Toronto with the vision of reducing homelessness in the city. This dataset is openly available to the public through the City of Toronto's Open Data Portal on an Open Government License - Toronto and was accessed through R using the R package `opendatatoronto` [@citeodt]. This dataset contains data from January 2020 to December 2021 and was last updated on January 7th, 2022.

## Methodology and Data Collection

The Shelter, Support & Housing Administration Division of the City of Toronto is the governmental service operator and manager of housing and homelessness services in Toronto. They offer services such as emergency shelters, street outreach, short-term respites and housing stability services such as drop-ins and eviction prevention [@citeshelter]. The Shelter System Flow data records the number of unique people who are entering and leaving the shelter system each month. In the data, the number of actively homeless is determined as the number people whom have used the shelter services within the past three months.

## Data Limitations

Since the data is collected from people accessing the homelessness services operated or managed by the City of Toronto, this data does not contain people who have been used these services such as people sleeping exclusively outdoors or using other homelessness services that are not using the Shelter Management Information System (SMIS). According to the dataset, based on a recent Street Needs Assessment (A city-wide count and survey of people experiencing homelessness in Toronto [@citesna]), there is an anticipated 18 per cent of people experiencing homelessness in Toronto that is not reflected in this dataset [@citedata].

There is a variety of reasons a person experiencing homelessness will not seek help from a shelter system. Example of this could be to avoid shelter restrictions such as curfews. Another potential reason, especially during the Covid-19 pandemic, is public health concerns [@citepallet]. This specific topic will be discussed further in section \@ref(covid).

## Data Characteristics

The original dataset contained 180 observations with 19 variables. These 180 observations are comprised of 7 categorical entries per month for 24 months (168 observations), and starting from January 2021, an additional entry was added for peoples identifying as Indigenous for a additional 12 observations (totaling 180).

Here is an extract of the dataset with a couple of variables of interest (Table \@ref(tab:extract)).

```{r extract, echo = FALSE, warning = FALSE, message = FALSE}
monthly_shelter_usage %>%
  select(`date(mmm-yy)`, population_group, actively_homeless, population_group_percentage) %>%
  slice(1:7) %>%
  kable(
    caption = "First ten rows of a dataset",
    col.names = c("Date (Month, Year)", "Population Group", "Actively Homeless", "Percentage of Group Population"),
    digits = 1,
    booktabs = TRUE,
    linesep=""
  )
```

Paragraph or two more about it

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}
# This needs to be about the random data that I sampled?
```

Talk way more about it.

# Discussion

## Covid-19 and its effects on homelessness shelter {#covid}

## Gender demographics

## Age demographics

## Weaknesses and next steps

Weaknesses and next steps should also be included.

# Conclusion

\newpage

# References